\section{Data Streams}
\frame{\sectionpage}

\begin{frame}{Stream Query Optimization}
    Modifying queries by changing graph topology and/ or operators to get better performance, as measured by
    \begin{enumerate}
        \item Throughput
        \item Latency
        \item Resource Usage.
    \end{enumerate}
    While preserving semantice of the original query.
\end{frame}

\begin{frame}{Stream Graph}
    Queries on Data streams can be thought of as directed graphs,
    \begin{enumerate}
    \item Edges represent streams and nodes represent operators.
    \item Root  and  leaf nodes are called sources and sinks, respectively.
    \end{enumerate}
    Whereas for traditional Databases, we have parse trees.
\end{frame}

\begin{frame}{Possible Optimizations}
    \begin{itemize}
        \item Batching
        \item Placement
        \item State sharing
        \item Load Balancing
        \item Algorithm selection
        \item Load Shedding
        \item Fusion
        \item Operator Separation
        \item Operator Reordering
        \item Redundancy elimination
        \item Fission
    \end{itemize}
\end{frame}

\begin{frame}{Operator Reordering}
    A reordering operator moves more selective operators, which reduce the data volumen upstream. This has the benefit of reducing the amount of data flowing into downstream compuration. Thus eliminatin unnecessary work. 
    We focus on finding the optimal method of executing associative opeators.
\end{frame}

\begin{frame}{Research Question}
    Is it possible to apply deep reinforcement learning to find the optimal ordering of executing associative operators?
\end{frame}
